%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Advanced Features}\label{chap:feat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\section{Supported kinds of FFT}

\subsection{c2c-FFT}
FFT of Complex inputs with complex outputs

\subsection{r2c-FFT and c2r-FFT}
FFT of real inputs with Hermitian symmetric complex outputs
\begin{compactitem}
  \item \code{local_size} functions give back number of complex numbers that must be allocated
  \item real array does not use padding\todo{no padding for r2c-PFFT}
  \item ghost cell send is based on r2r and c2c ghost cell send with appropriate parameters \todo{example for r2c ghost cell send}
  \begin{compactitem}
    \item use physical size for c2c ghost cell send
    \item use \code{2*alloc_local} in \code{local_size_gc} function for r2r ghost cell send
    \item FFTW r2c computes \code{FFTW_FORWARD} and c2r computes \code{FFTW_BACKWARD} only~\cite{fftw-rdft}. In contrast, PFFT adds \code{sign} also for r2c/c2r transforms.
          This is implemented with a conjugation of the complex values.
  \end{compactitem}
\end{compactitem}




\subsection{r2r-FFT}
FFT of real inputs with real outputs



\begin{compactitem}
  \item how to do in place transforms
\end{compactitem}


\section{Reduction of communication by transposed data layouts}
By providing \code{pfft\_flags} the user can choose between three different data layouts of the output
array. Table~\ref{tab:pfft_flags} shows the flags and their corresponding decompositions.
\begin{table}[h]
  \begin{tabular}{lll}
    flag & input layout & output layout \\
    \code{PFFT\_DEFAULT}     & n0/P0 x n1/P1 x n2 & n0/P1 x n1/P0 x n2 \\
    \code{PFFT\_TRANSPOSED}  & n0/P0 x n1/P1 x n2 & n2/P1 x n1/P0 x n0 \\
    \code{PFFT\_P3DFFT}      & n0/P0 x n1/P1 x n2 & n2/P0 x n0/P1 x n1
  \end{tabular}
  \caption{Impact of predefined PFFT flags on the parallel data layout.}
  \label{tab:pfft_flags}
\end{table}


Assume a three dimensional input function \code{fftw\_complex input(ptrdiff\_t k0, ptrdiff\_t k1, ptrdiff\_t k2);}
Data init:
\begin{lstlisting}
int m=0;
for(int k0=0; k0<local_ni[0]; k0++)
  for(int k1=0; k1<local_ni[1]; k1++)
    for(int k2=0; k2<local_ni[2]; k2++)
      in[m++] = input(local_i_start[0]+k0, local_i_start[1]+k1, local_i_start[2]+k2);
\end{lstlisting}


%------------------------------------------------------------------------------
\section{How to Deal with FFT Index Shifts in Parallel}
%------------------------------------------------------------------------------
Let $n\in2\N$. A common problem is that the index of the FFT input and/or output array runs between $-\nicefrac n2,\hdots,\nicefrac n2-1$,
but the FFT library requires them to run between $0,\hdots,n-1$. With serial program execution one can easily remap the input data $\hat g_k$
in a way that is suitable for the library, i.e.,
\begin{equation*}
  \hat f_k := \hat g_{(k-\nicefrac n2\bmod n)}, \quad k = 0,\hdots,n-1.
\end{equation*}
Similarly, one could remap the outputs of the library $f_l$, $l=0,\cdots,n-1$ in the opposite direction in order to get the
required outputs, i.e.,
\begin{equation*}
  g_l := f_{l \bmod n}, \quad l = -\nicefrac n2,\hdots,\nicefrac n2-1.
\end{equation*}
These shifts are also known as \code{fftshift} in Matlab.

However, with distributed memory these \code{fftshift} operations require more complex data movements and result in a global communication.
For example, the first index of the array moves to the middle and, therefore, the corresponding data move to another MPI process.
Fortunately, this communication can be avoided at the cost of little extra computation.
At the end of the section we present two PFFT library functions that perform the necessary pre- and postprocessing
for shifted input and output index sets.

\subsection{Shift with half the FFT size}

The special case of input shift $k_s=-\nicefrac n2$ and/or output shift $l_s=-\nicefrac n2$ is supported by PFFT.
User can choose to shift the input (\verb+PFFT_SHIFTED_IN+) and/or to shift the output (\verb+PFFT_SHIFTED_OUT+).
\todo{this flag can be used for \code{local_size} and planning}

Here, we are interested in the computation of
\begin{equation*}
  g_l = \sum_{k=-\nicefrac{n_i}{2}}^{\nicefrac{n_i}{2}-1} \hat g_k \eim{kl/n}, \quad l=-\nicefrac{n_o}{2},\hdots,\nicefrac{n_o}{2}-1  
\end{equation*}
with $n, n_i, n_o \in 2\N$ and $n>n_i$, $n>n_o$.

With an index shift of $\nicefrac n2$ both in $k$ and $l$ this equivalent to the computation of
\begin{align*}
  g_{(l-\nicefrac{n}{2})}
  &= \sum_{k=\nicefrac{n}{2}-\nicefrac{n_i}{2}}^{\nicefrac{n}{2}+\nicefrac{n_i}{2}-1}
     \hat g_{(k-\nicefrac{n}{2})} \eim{(k-\nicefrac n2)(l-\nicefrac n2)/n} \\
  &= \e^{+\pi\ti l} 
       \sum_{k=\nicefrac{n}{2}-\nicefrac{n_i}{2}}^{\nicefrac{n}{2}+\nicefrac{n_i}{2}-1}
       \left(\hat g_{(k-\nicefrac{n}{2})}\e^{+\pi\ti (k-\nicefrac n2)}\right) \eim{kl/n} \\
  &= \e^{+\pi\ti(l-\nicefrac n2)} 
     \underbrace{
       \sum_{k=\nicefrac{n}{2}-\nicefrac{n_i}{2}}^{\nicefrac{n}{2}+\nicefrac{n_i}{2}-1}
       \underbrace{\left(\hat g_{(k-\nicefrac{n}{2})}\e^{+\pi\ti k}\right)}_{\hat f_k} \eim{kl/n}
     }_{f_l}
\end{align*}
for $ l=\nicefrac n2-\nicefrac{n_o}{2},\hdots,\nicefrac n2 +\nicefrac{n_o}{2}-1$.
Therefore, we get the following algorithm

\begin{equation*}
  f_l = \sum_{k=0}^n \hat g_k \eim{kl/n}, \quad l=-\nicefrac{n_o}{2},\hdots,\nicefrac{n_o}{2}-1  
\end{equation*}

The special case $k_s=-\frac{n_i}{2}, l_s=-\frac{n_o}{2}$ corresponds to the shifts the arrays (\textsf{FFTSHIFT})
\begin{algorithm}
  \begin{algorithmic}[1]
    \itemsep=1.1ex
    \State For $k=0,\hdots,n-1$ set $\hat f_k = 0$.
    \State For $k=-\nicefrac{n_i}{2},\hdots,\nicefrac{n_i}{2}-1$ compute $\hat f_{(k+\nicefrac{n}{2})} = (-1)^{(k+\nicefrac{n}{2})} \hat g_{k}$.
    \State For $l=0,\hdots,n-1$ compute $f_l = \sum_{k=0}^{n} \hat f_k \eim{kl/n}$ using PFFT.
    \State For $l=-\nicefrac{n_o}{2},\hdots,\nicefrac{n_o}{2}-1$ compute $g_l = (-1)^l f_{(l+n/2)} $.
  \end{algorithmic}
\end{algorithm}


Note, that this shift implies that the library deals with pruned FFTs in a special way, i.e., half of the zeros are added
at the beginning of the inputs and the other half is added at the end.






\subsection{Arbitrary shifts}
More general shifts must be done by the user.


In a more general setting, we are interested in the computation of FFTs with shifted index sets, i.e., assume $k_s,l_s\in\Z$ and compute
\begin{equation*}
  g_l = \sum_{k=k_s}^{n_i+k_s-1} \hat g_k \eim{kl/n},
  \quad l=l_s,\hdots,n_o+l_s-1\,.
\end{equation*}
Because of the periodicity of the FFT this can be easily performed by \algname~\ref{alg:fftshift_translation}.
\begin{algorithm}\label{alg:fftshift_translation}
  \begin{algorithmic}[1]
    \itemsep=1.1ex
    \State For $k=0,\hdots,n_i-1$ assign $\hat f_k = \hat g_{(k+k_s\bmod n_i)}$.
    \State For $l=0,\hdots,n_o-1$ compute $f_l = \sum_{k=0}^{n_i} \hat f_k \eim{kl/n}$ using PFFT.
    \State For $l=0,\hdots,n_o-1$ assign $g_l = f_{(l-l_s\bmod n_o)}$.
  \end{algorithmic}
  \caption{Shifted FFT with explicit data movement.}
\end{algorithm}
However, this involves explicit data movement since the sequence of data changes.
For a our parallel data decomposition the change of data layout requires data communication.
A simple index shift results in the computation of
\begin{align*}
  g_{l+l_s}
  &=
    \sum_{k=k_s}^{n_i+k_s-1} \hat g_k \eim{k(l+l_s)/n}
    =
    \sum_{k=0}^{n_i-1} \hat g_{k+k_s} \eim{(k+k_s)(l+l_s)/n} \\
  &=
    \eim{k_sl/n} \sum_{k=0}^{n_i-1} \underbrace{\left(\hat g_{k+k_s}\eim{(k+k_s)l_s/n}\right)}_{=: \hat f_k} \eim{kl/n}
\end{align*}
for all $l=0,\hdots,n_o-1$. The resulting \algname~\ref{alg:fftshift_modulation} preserves the sequence of
data at the price of some extra computation.
\begin{algorithm}\label{alg:fftshift_modulation}
  \begin{algorithmic}[1]
    \itemsep=1.1ex
    \State For $k=0,\hdots,n_i-1$ compute $\hat f_k = \hat g_{(k+k_s)} \eim{(k+k_s)l_s/n}$.
    \State For $l=0,\hdots,n_o-1$ compute $f_l = \sum_{k=0}^{n_i} \hat f_k \eim{kl/n}$ using PFFT.
    \State For $l=0,\hdots,n_o-1$ compute $g_{(l+l_s)} = f_l \eim{k_sl/n}$.
  \end{algorithmic}
  \caption{Shifted FFT without explicit data movement.}
\end{algorithm}

The special case $k_s=-\frac{n_i}{2}, l_s=-\frac{n_o}{2}$ corresponds to the shifts the arrays (\textsf{FFTSHIFT})
\begin{algorithm}
  \begin{algorithmic}[1]
    \itemsep=1.1ex
    \State For $k=0,\hdots,n_i-1$ compute $\hat f_k = \hat g_{(k-\nicefrac{n_i}{2})} \e^{+\pi\ti (k-\nicefrac{n_i}{2})n_o/n}$.
    \State For $l=0,\hdots,n_o-1$ compute $f_l = \sum_{k=0}^{n_i} \hat f_k \eim{kl/n}$ using PFFT.
    \State For $l=0,\hdots,n_o-1$ compute $g_{(l-\nicefrac{n_o}{2})} = f_l \e^{+\pi\ti n_i l/n}$.
  \end{algorithmic}
\end{algorithm}




%------------------------------------------------------------------------------
\section{Parallel pruned FFT}
%------------------------------------------------------------------------------
Within PFFT we define a pruned FFT as
\begin{equation*}
  g_l = \sum_{k=0}^{n_i-1} \hat g_{k} \eim{kl/n}, \quad l=0,\hdots,n_o-1.
\end{equation*}
Formally, this is equivallent to the following regular size $n$ FFT
\begin{equation*}
  f_l = \sum_{k=0}^{n-1} \hat f_{k} \eim{kl/n}, \quad l=0,\hdots,n,
\end{equation*}
with 
\begin{equation*}
  \hat g_k := 
  \begin{cases}
  \hat f_k, &: k=0,\hdots,n_1-1, \\
  0         &: k=n_i,\hdots,n-1,    
  \end{cases}
\end{equation*}
and $f_l := g_l$, $k=0,\hdots,n_o-1$. I.e., we add $n-n_i$ zeros at the end of the input array and throw away $n-n_o$ entries at the end of the output array.


The definition of pruned FFT changes for \code{PFFT_SHIFTED_IN} and \code{PFFT_SHIFTED_OUT}.


\section{Skip FFTs along some dimensions}

\code{int skip_trafos[rnk_n]}, but skip of last \code{rnk_n - rnk_pm + 2} dims is not yet implemented (needs other call to serial multidimensional FFT). 





%------------------------------------------------------------------------------
\section{Computation of block distribution}
%------------------------------------------------------------------------------
\subsection{One-dimensional distribution}
%------------------------------------------------------------------------------
Assume $N$ grid points to be distributed on $P$ MPI processes. The block size $B$ is given by
\begin{equation*}
  B = \left\lceil \frac{N}{P} \right\rceil\,.
\end{equation*}
The first $R=\left\lfloor\frac{P}{B}\right\rfloor$ processes get $B$ grid points. The next process gets $N-BR$ grid points and all remaining processes get nothing.


\subsection{Remap three- into two-dimensional distribution}
%------------------------------------------------------------------------------
\begin{equation*}
  \frac{n_0}{P_0} \times \frac{n_1}{P_1} \times \frac{n_2}{Q_0Q_1}
  \to
  \frac{n_0}{P_0} \times \frac{n_1}{P_1} \times \frac{n_2}{Q_0Q_1}
  \to
  \frac{n_0}{P_0Q_0} \times \frac{n_1}{P_1Q_1} \times n_2
\end{equation*}

The algorithm work as follows:
\begin{align*}
  \frac{n_0}{P_0} \times \frac{n_1}{P_1} \times \frac{n_2}{Q_0Q_1}
  &\to
  \frac{n_2}{Q_0Q_1} \times \frac{n_0}{P_0} \times \frac{n_1}{P_1}
  &\to
  \frac{n_2}{Q_0} \times \frac{n_0}{P_0} \times \frac{n_1}{P_1Q_1} \\
  &\to
  n_2 \times \frac{n_0}{P_0Q_0} \times \frac{n_1}{P_1Q_1}
  &\to
  \frac{n_0}{P_0Q_0} \times \frac{n_1}{P_1Q_1} \times n_2
\end{align*}

\subsubsection{Example: $\mathbf n=(674, 674, 674)^\top$, $\mathbf P = (16,16,8)$ with normal 3d decomposition}
$\frac{674}{16}$:
\begin{itemize}
 \item $B = \left\lceil \frac{674}{16} \right\rceil = 43$
 \item $R = \left\lfloor \frac{674}{43} \right\rfloor = 15$
 \item $674 - R*B = 29$ 
\end{itemize}
\begin{equation*}
  \frac{674}{16} = \left[15\times 43, 1\times 29 \right]
\end{equation*}

$\frac{674}{8}$:
\begin{itemize}
 \item $B = \left\lceil \frac{674}{16} \right\rceil = 85$
 \item $R = \left\lfloor \frac{674}{43} \right\rfloor = 7$
 \item $674 - R*B = 79$
\end{itemize}
\begin{equation*}
  \frac{674}{16} = \left[7\times 85, 1\times 79 \right]
\end{equation*}

\subsubsection{Example: $\mathbf n=(674, 674, 674)^\top$, $\mathbf P = (16,16,8)$ with 3d decomposition suitable for 3d to 2d remap}

\begin{lstlisting}
/* n0/(p0*q0) x n1/(p1*q1) x n2 */
oblk[0] = PX(global_block_size)(n[0], PFFT_DEFAULT_BLOCK, p0*q0);
oblk[1] = PX(global_block_size)(n[1], PFFT_DEFAULT_BLOCK, p1*q1);
oblk[2] = n[2];

/* n0/p0 x n1/p1 x n2/(q0*q1) */
iblk[0] = oblk[0]*q0;
iblk[1] = oblk[1]*q1;
iblk[2] = PX(global_block_size)(n[2], PFFT_DEFAULT_BLOCK, q0*q1);
//PX(global_block_size)(blk1[2], PFFT_DEFAULT_BLOCK, q1);

/* n0/p0 x n1/(p1*q1) x n2/q0 */
mblk[0] = oblk[0]*q0;
mblk[1] = oblk[1];
mblk[2] = iblk[2]*q1;
\end{lstlisting}

Compute the three-dimensional block in the following way:
\begin{enumerate}
 \item compute blocks $B_0^\textrm{2d}, B_1^\textrm{2d}$ for 2d decomposition
 \item compute blocks for 3d decomposition via $B_0^\textrm{3d} = B_0^\textrm{2d} Q_0$, $B_1^\textrm{3d} = B_1^\textrm{2d} Q_1$ and the default block size $B_2^\textrm{3d} = \left\lceil\frac{n_2}{Q_0Q_1}\right\rceil$
\end{enumerate}


$P_0 = 16, P_1 = 16, Q_0=4, Q_1=2$

$\frac{n_0}{P_0Q_0} = \frac{674}{16\cdot 4}$:
\begin{itemize}
 \item $B_0^\textrm{2d} = \left\lceil \frac{674}{64} \right\rceil = 11$
 \item $B_0^\textrm{3d} = B_0^\textrm{2d}Q_0 = 44$
 \item $R_0^\textrm{3d} = \left\lfloor \frac{674}{44} \right\rfloor = 15$
 \item $674 - R_0^\textrm{3d}*B_0^\textrm{3d} = 15$
\end{itemize}
\begin{equation*}
  \frac{n_0}{P_0} = \frac{674}{16} = \left[15\times 44, 1\times 14 \right]
\end{equation*}
$\frac{n_1}{P_1Q_1} = \frac{674}{16\cdot 2}$:
\begin{itemize}
 \item $B_1^\textrm{2d} = \left\lceil \frac{674}{32} \right\rceil = 22$
 \item $B_1^\textrm{3d} = B_1^\textrm{2d}Q_1 = 44$
 \item $R_1^\textrm{3d} = \left\lfloor \frac{674}{44} \right\rfloor = 15$
 \item $674 - R_1^\textrm{3d}*B_1^\textrm{3d} = 15$
\end{itemize}
\begin{equation*}
  \frac{n_1}{P_1} = \frac{674}{16} = \left[15\times 44, 1\times 14 \right]
\end{equation*}

\begin{itemize}
 \item $B_2^\textrm{3d} = \left\lceil \frac{674}{8} \right\rceil = 85$
 \item $R_2^\textrm{3d} = \left\lfloor \frac{674}{85} \right\rfloor = 7$
 \item $674 - R_2^\textrm{3d}*B_2^\textrm{3d} = 79$
\end{itemize}
\begin{equation*}
  \frac{n_2}{Q_0Q_1} = \frac{674}{8} = \left[7\times 85, 1\times 79 \right]
\end{equation*}

%------------------------------------------------------------------------------
\section{Ghost cell support}
%------------------------------------------------------------------------------


%------------------------------------------------------------------------------
\section{Single and long double precision}
%------------------------------------------------------------------------------


%------------------------------------------------------------------------------
\section{Local block functions}
%------------------------------------------------------------------------------


